---
title: "R Notebook"
output: html_notebook
---


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(janitor)
```

```{r}
`%!in%` <- negate(`%in%`)

pal <- c("Male" = "#a6bddb", "Female" = "#fa9fb5")

nhs <- c( "#ee9c00", "#ff0000", 
         "#b80068", "#00a15f",
         "#0391bf", "#67bf29", "#00684d")

hb <- c('#a6cee3','#1f78b4','#b2df8a',
         '#33a02c','#fb9a99','#e31a1c',
         '#fdbf6f','#ff7f00','#cab2d6',
         '#6a3d9a','#525252')
```


```{r}
council_areas <- read_csv(
  "../raw_data/dz2001_codes_and_labels_21042020.csv") %>% 
  clean_names()
```


```{r}
areas <- council_areas %>% 
  select(contains("ca")) %>% 
  rename(area_codes = "ca",
        council_name = "ca_name") %>% 
  add_row(area_codes = "S92000003", council_name = "Scotland") %>% 
  add_row(area_codes = "S12000046", council_name = "Glasgow City") %>%
  add_row(area_codes = "S12000044", council_name = "North Lanarkshire") %>%
  group_by(area_codes, council_name) %>% 
  unique() %>% 
  ungroup()

areas
```

```{r}
health_board <- council_areas %>% 
  select(contains("hb")) %>% 
  rename(health_board = "hb") %>% 
  add_row(health_board = "S92000003", hb_name = "Scotland") %>%
  add_row(health_board = "S08000018", hb_name = "NHS Fife") %>%
  add_row(health_board = "S08000021", 
          hb_name = "NHS Greater Glasgow and Clyde") %>%
  add_row(health_board = "S08000023", hb_name = "NHS Lanarkshire") %>%
  group_by(health_board, hb_name) %>% 
  unique() %>% 
  ungroup()

health_board
```
```{r}
hb_stroke <- read_csv("../raw_data/stroke_activitybyhbr.csv") %>% 
  clean_names() %>% 
  rename(health_board = "hbr")
```


```{r}
health_board_stroke <- right_join(health_board, 
                                  hb_stroke, 
                                  by = "health_board") %>% 
  select(hb_name, financial_year, admission_type, age_group, sex, 
         diagnosis, number_of_discharges, crude_rate, easr) %>%
  mutate(financial_year = str_sub(financial_year, start = 1, end = 4),
         financial_year = as.numeric(financial_year)) %>% 
  mutate(age_group = factor(age_group, 
                            levels = c("All", "0-44 years", 
                                       "45-64 years", "65-74 years", 
                                       "under75 years", "75plus years" )))
   # filter(is.na(hb_name))

health_board_stroke
```

# What is the most common stroke diagnosis in Scotland?

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "Scotland",
          across(admission_type:sex, ~.x=="All")) %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges, 
             col = diagnosis)) +
  geom_point() +
  geom_line(aes(group = diagnosis)) +
  scale_colour_manual(values = nhs) +
  labs(title = "Common Stroke Diagnosis in Scotland \n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Types of Strokes\n") +
  theme(plot.title = element_text(hjust = 0.5))
```


# Differences in stroke diagnoses between different demographics

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "Scotland",
          across(admission_type:age_group, ~.x=="All"),
         sex %!in% "All") %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges, 
             col = sex)) +
  geom_point() +
  geom_line(aes(group = sex)) +
  scale_colour_manual(values = pal) +
  labs(title = "Common Stroke Diagnosis in Scotland by Gender\n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Strokes by Gender\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "Scotland",
          across(age_group:sex, ~.x=="All"),
         admission_type %!in% "All") %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges, 
             col = admission_type)) +
  geom_point() +
  geom_line(aes(group = admission_type)) +
  scale_colour_manual(values = nhs) +
  labs(title = "Common Stroke Diagnosis in Scotland by Admission Types\n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Admission Types\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "Scotland",
         sex %in% "All",
         admission_type %in% "All",
         age_group %!in% c("All", "under75 years")) %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges, 
             col = age_group)) +
  geom_point() +
  geom_line(aes(group = age_group)) +
  scale_colour_manual(values = nhs) +
  labs(title = "Common Stroke Diagnosis in Scotland by Age Group\n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Age Group\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```
```{r}
unique(health_board_stroke$hb_name)
```


```{r}
health_board_stroke %>% 
  filter(hb_name %!in% c("Scotland", "NHS Western Isles",
                         "NHS Orkney", "NHS Shetland"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %!in% c("Subarachnoid Haemorrhage", 
                           "TIAs and related syndromes")) %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges,
             group = hb_name, col = hb_name )) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = hb,
                      labels = c("Ayrshire", "Borders", "Dumfries &\n Galloway",
                                 "Fife", "Forth Valley", "Grampian",
                                 "Greater\n Glasgow", "Highland", "Lanarkshire",
                                 "Lothian", "Tayside")) +
  labs(title = "Common Stroke Diagnosis by Health Boards \n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Health Boards\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

Omitted the NHS Western Isles, NHS Orkney, NHS Shetland, Subarachnoid Haemorrhage, TIAs and related syndromes doe to very small numbers

```{r}
health_board_stroke %>% 
  filter(hb_name %!in% c("Scotland", "NHS Western Isles",
                         "NHS Orkney", "NHS Shetland"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges,
             group = hb_name, col = hb_name )) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = hb, guide = "none") +
  geom_text(aes(label = "Borders", 
                x = 2019, y = 600), color = "#1f78b4") +
  geom_text(aes(label = "Greater \nGlasgow", 
                x = 2019, y = 8000), color = "#fdbf6f") +
  geom_text(aes(label = "Dumfries", 
                x = 2019, y = 1000), color = "#b2df8a") +
  geom_text(aes(label = "Lothian", 
                x = 2019, y = 4200), color = "#6a3d9a") +
  geom_text(aes(label = "Lanarkshire", 
                x = 2019, y = 5000), color = "#cab2d6") +
  geom_text(aes(label = "Forth Valley", 
                x = 2019, y = 1400), color = "#fb9a99") +
  geom_text(aes(label = "Grampian ", 
                x = 2019, y = 1750), color = "#e31a1c") +
  geom_text(aes(label = "Highland", 
                x = 2019, y = 2100), color = "#ff7f00") +
  geom_text(aes(label = "Fife", 
                x = 2019, y = 2550), color = "#33a02c") +
  geom_text(aes(label = "Tayside", 
                x = 2019, y = 2900), color = "#525252") +
  geom_text(aes(label = "Ayrshire", 
                x = 2019, y = 3300), color = "#a6cee3") +
   scale_x_continuous(limits = c(2009, 2019.5), 
                      breaks = c(2010, 2012, 2014, 2016, 2018)) +
  labs(title = "Cerebrovascular Disease Diagnosis by Health Boards \n",
       x = "Financial Year",
       y = "Number of Discharges \n") +
  theme(plot.title = element_text(hjust = 0.5))
```





  "NHS Grampian" : Aberdeen, Aberdeenshire and Moray. 
  
 "NHS Tayside" :  Angus, the City of Dundee and Perth and Kinross.
 
 NHS Forth Valley : Clackmannanshire, Falkirk and Stirling 
 
"NHS Ayrshire and Arran" : North Ayrshire South Ayrshire East Ayrshire

 "NHS Lothian" : City of Edinburgh, East Lothian, Midlothian and West Lothian 
 
 NHS Greater Glasgow and Clyde : City of Glasgow, East Dunbartonshire, East Renfrewshire, Inverclyde, Renfrewshire and West Dunbartonshire. 
 
 NHS Lanarkshire : North Lanarkshire and South Lanarkshire



# Differences in number of cases by health board vs by local authority

```{r}
ca_stroke <- read_csv("../raw_data/stroke_activitybyca.csv") %>% 
  clean_names() %>% 
  rename(area_codes = "ca")
```

```{r}
council_stroke <- right_join(areas, ca_stroke, by = "area_codes") %>% 
  select(council_name, financial_year, admission_type, age_group, 
         sex, diagnosis, number_of_discharges, crude_rate, easr) %>%
  mutate(financial_year = str_sub(financial_year, start = 1, end = 4),
         financial_year = as.numeric(financial_year)) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(age_group = factor(age_group, 
                            levels = c("All", "0-44 years", 
                                       "45-64 years", "65-74 years", 
                                       "under75 years", "75plus years" )))

   # filter(is.na(council_name))

council_stroke
```



```{r}
grampian <- health_board_stroke %>% 
  filter(hb_name %in% "NHS Grampian",
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  rename(council_name = hb_name)


nhs_grampian <-  bind_rows(council_stroke, grampian)

nhs_grampian %>% 
   filter(council_name %in% c("Aberdeen City", 
                              "Aberdeenshire", "Moray", 
                              "NHS Grampian"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges)) +
  geom_point(aes(group = council_name, col = council_name)) +
  geom_line(aes(group = council_name, col = council_name)) +
    scale_colour_manual(values = nhs, 
                      labels = c("Aberdeen City", 
                                 "Aberdeenshire", "Moray", 
                                 "NHS Grampian")) +
  labs(title = "Cerebrovascular Disease in NHS Grampian \n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Local Authority\n") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
nhs_grampian %>% 
   filter(council_name %in% c("Aberdeen City", 
                        "Aberdeenshire", "Moray"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```


```{r}
health_board_stroke %>% 
  filter(hb_name %in% "NHS Grampian",
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  select(hb_name, financial_year, number_of_discharges) 
```

For Cerebrovascular Disease NHS Grampian has an extra 6, 0.3% in 2018/19 compared to the local authority's 

```{r}
nhs_grampian %>% 
   filter(council_name %in% c("Aberdeen City", 
                        "Aberdeenshire", "Moray"),
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```


```{r}
health_board_stroke %>% 
  filter(hb_name %in% "NHS Grampian",
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```

NHS Grampian across all the four types of strokes has an extra 8, 0.2% in 2016/17 and an extra 6, 0.2% in 2018/19 compared to the local authority's


```{r}
lothian <- health_board_stroke %>% 
  filter(hb_name %in% "NHS Lothian",
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  rename(council_name = hb_name)


nhs_lothian <-  bind_rows(council_stroke, lothian)

nhs_lothian %>% 
   filter(council_name %in% c("City of Edinburgh", 
                        "East Lothian", "Midlothian", 
                        "West Lothian", "NHS Lothian"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  ggplot(aes(x = financial_year, 
             y = number_of_discharges)) +
  geom_point(aes(group = council_name, col = council_name)) +
  geom_line(aes(group = council_name, col = council_name)) +
  scale_colour_manual(values = nhs, 
                      labels = c("City of \nEdinburgh", 
                                 "East Lothian", "Midlothian",
                                 "NHS Lothian", "West Lothian")) +
  labs(title = "Cerebrovascular Disease in NHS Lothian \n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Local Authority\n") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
nhs_lothian %>% 
   filter(council_name %in% c("City of Edinburgh", 
                        "East Lothian", "Midlothian", 
                        "West Lothian"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```


```{r}
health_board_stroke %>% 
  filter(hb_name %in% "NHS Lothian",
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  select(hb_name, financial_year, number_of_discharges) 
```

For Cerebrovascular Disease NHS Lothian has an extra 5, 0.1% in 2017/18 and an extra 24, 0.6% in 2018/19 compared to the local authority's


```{r}
nhs_lothian %>% 
   filter(council_name %in% c("City of Edinburgh", 
                        "East Lothian", "Midlothian", 
                        "West Lothian"),
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "NHS Lothian",
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges)) 
```
NHS Lothian across all the four types of strokes has an extra 8, 0.1% in 2017/18 and an extra 45, 0.6% in 2018/19 compared to the local authority's



```{r}
glasgow <- health_board_stroke %>% 
  filter(hb_name %in% "NHS Greater Glasgow and Clyde",
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>% 
  rename(council_name = hb_name)


nhs_glasgow <-  bind_rows(council_stroke, glasgow)

nhs_glasgow %>% 
   filter(council_name %in% c("Glasgow City", 
                        "East Dunbartonshire", "East Renfrewshire", 
                        "Inverclyde", "Renfrewshire", "West Dunbartonshire",
                        "NHS Greater Glasgow and Clyde"),
         across(admission_type:sex, ~.x=="All"),
         diagnosis %in% "Cerebrovascular Disease") %>%
  ggplot(aes(x = financial_year, 
             y = number_of_discharges)) +
  geom_point(aes(group = council_name, col = council_name)) +
  geom_line(aes(group = council_name, col = council_name)) +
  scale_colour_manual(values = nhs, 
                      labels = c("East \nDunbartonshire", "East \nRenfrewshire",
                                 "Glasgow City", "Inverclyde",
                                 "NHS Greater \nGlasgow ",
                                 "Renfrewshire", "West \nDunbartonshire")) +
  labs(title = "Cerebrovascular Disease in NHS Greater Glasgow \n",
       x = "Financial Year",
       y = "Number of Discharges \n",
       colour = "Local Authority\n") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
nhs_glasgow %>% 
   filter(council_name %in% c("Glasgow City", 
                        "East Dunbartonshire", "East Renfrewshire", 
                        "Inverclyde", "Renfrewshire", "West Dunbartonshire"),
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges))
```

```{r}
health_board_stroke %>% 
  filter(hb_name %in% "NHS Greater Glasgow and Clyde",
         across(admission_type:sex, ~.x=="All")) %>% 
  group_by(financial_year) %>% 
  summarise(sum_of_discharges = sum(number_of_discharges)) 
```

NHS Greater Glasgow and Clyde across all the four types of strokes has an extra 3, 0.02% in 2016/17, an extra 14, 0.08% in 2017/18 and an extra 86, 0.5% in 2018/19 compared to the local authority's





```{r}
 # health_board_stroke %>% 
 #  filter(across(admission_type:sex, ~.x=="All")) %>% 
 #  group_by(hb_name, financial_year) %>% 
 #  summarise(sum_hb_discharges = sum(number_of_discharges))
```

```{r}
# council_stroke %>% 
#   filter(across(admission_type:sex, ~.x=="All")) %>% 
#   group_by(hb_name, financial_year) %>% 
#   summarise(sum_hb_discharges = sum(number_of_discharges))
```





```{r}
unique(nhs_grampian$council_name)
```








# Differences in mortality rates for different types of stroke

```{r}
stroke_mortality_hb <- read_csv("../raw_data/stroke_mortalitybyhbr.csv") %>% 
  clean_names() %>% 
  rename(health_board = "hbr")
```

```{r}
health_board_mortality <- right_join(health_board, 
                                     stroke_mortality_hb,
                                     by = "health_board") %>% 
  select(hb_name, year, age_group, sex, 
         diagnosis, number_of_deaths, crude_rate, easr) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(age_group = factor(age_group, 
                            levels = c("All", "0-44 years", 
                                       "45-64 years", "65-74 years", 
                                       "under75 years", "75plus years" )))
  
   # filter(is.na(hb_name))

health_board_mortality
```

```{r}
unique(health_board_mortality$age_group)
```


```{r}
health_board_mortality %>% 
  filter(hb_name %in% "Scotland",
          across(age_group:sex, ~.x=="All")) %>% 
  ggplot(aes(x = year, 
             y = number_of_deaths, 
             col = diagnosis)) +
  geom_point() +
  geom_line(aes(group = diagnosis)) +
  scale_colour_manual(values = nhs, 
                      labels = c("Cerebrovascular \nDisease", "Stroke",
                                 "Subarachnoid \nHaemorrhage")) +
  labs(title = "Stroke Mortality Rate in Scotland \n",
       x = "Financial Year",
       y = "Mortality Rate \n",
       colour = "Types of Strokes\n") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

Again Subarachnoid Haemorrhage, TIAs and related syndromes very small numbers so will leave out.


```{r}
health_board_mortality %>% 
  filter(hb_name %in% "Scotland",
          across(age_group, ~.x=="All"),
         sex %!in% "All",
         diagnosis %!in% c("Subarachnoid Haemorrhage", 
                           "TIAs and related syndromes")) %>% 
  ggplot(aes(x = year, 
             y = number_of_deaths, 
             col = sex)) +
  geom_point() +
  geom_line(aes(group = sex)) +
  scale_colour_manual(values = pal) +
  labs(title = "Stroke Mortality Rate in Scotland by Gender\n",
       x = "Financial Year",
       y = "Mortality Rate \n",
       colour = "Strokes by Gender\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

```{r}
health_board_mortality %>% 
  filter(hb_name %in% "Scotland",
         sex %in% "All",
         age_group %!in% c("All", "under75 years"),
         diagnosis %!in% c("Subarachnoid Haemorrhage", 
                           "TIAs and related syndromes")) %>% 
  ggplot(aes(x = year, 
             y = number_of_deaths, 
             col = age_group)) +
  geom_point() +
  geom_line(aes(group = age_group)) +
  scale_colour_manual(values = nhs) +
  labs(title = "Stroke Mortality Rate in Scotland by Age Group\n",
       x = "Financial Year",
       y = "Mortality Rate \n",
       colour = "Age Group\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

```{r}
health_board_mortality %>% 
  filter(hb_name %!in% c("Scotland", "NHS Western Isles",
                         "NHS Orkney", "NHS Shetland"),
         across(age_group:sex, ~.x=="All"),
         diagnosis %!in% c("Subarachnoid Haemorrhage", 
                           "TIAs and related syndromes")) %>% 
  ggplot(aes(x = year, 
             y = number_of_deaths,
             group = hb_name, col = hb_name )) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = hb,
                      labels = c("Ayrshire", "Borders", "Dumfries &\n Galloway",
                                 "Fife", "Forth Valley", "Grampian",
                                 "Greater\n Glasgow", "Highland", "Lanarkshire",
                                 "Lothian", "Tayside")) +
  labs(title = "Stroke Mortality Rate by Health Boards \n",
       x = "Financial Year",
       y = "Mortality Rate \n",
       colour = "Health Boards\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

```{r}
health_board_mortality %>% 
  filter(hb_name %!in% c("Scotland", "NHS Western Isles",
                         "NHS Orkney", "NHS Shetland"),
         across(age_group:sex, ~.x=="All"),
         diagnosis %!in% c("Subarachnoid Haemorrhage", 
                           "TIAs and related syndromes")) %>% 
  ggplot(aes(x = year, 
             y = crude_rate,
             group = hb_name, col = hb_name )) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = hb,
                      labels = c("Ayrshire", "Borders", "Dumfries &\n Galloway",
                                 "Fife", "Forth Valley", "Grampian",
                                 "Greater\n Glasgow", "Highland", "Lanarkshire",
                                 "Lothian", "Tayside")) +
  labs(title = "Stroke Crude Rate by Health Boards \n",
       x = "Financial Year",
       y = "Crude Rate \n",
       colour = "Health Boards\n") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~diagnosis)
```

The crude rate per 100,000 population.


NHS Grampian	2009	0-44 years	Male	Cerebrovascular Disease	1	0.615127	0.608219

# Predict upcoming diagnoses/mortality rates for future years

```{r}
clean_health_board_mortality <- health_board_mortality %>% 
  filter(across(age_group:sex, ~.x!="All"),
         age_group %!in% "under75 years") %>% 
  select(-easr) %>% 
  na.omit()

clean_health_board_mortality
```

Omitted the NHS Western Isles, NHS Orkney, NHS Shetland as they had no data on deaths.

```{r}
summary(clean_health_board_mortality)
```


```{r}
glimpse(clean_health_board_mortality)
```

```{r}
n_data <- nrow(clean_health_board_mortality)

test_index <- sample(1:n_data, size = n_data*0.2)

mortality_test <- slice(clean_health_board_mortality, test_index)

mortality_train <- slice(clean_health_board_mortality, -test_index)
```

```{r}
mortality_test %>% 
  tabyl(number_of_deaths)
```

```{r}
mortality_train %>% 
  tabyl(number_of_deaths)
```

```{r}
library(rpart)
library(rpart.plot)

mortality_fit <- rpart(
  formula = number_of_deaths ~ ., 
  data = mortality_train, 
  method = 'class'
)

rpart.plot(mortality_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)
```



```{r}
rpart.rules(mortality_fit, cover = TRUE)
```




```{r}
stroke_mortality_ca <- read_csv("../raw_data/stroke_mortalitybyca.csv") %>% 
  clean_names()
```



```{r}
stroke_mortalitybyca %>% 
  
```

```{r}
hb_stroke <- read_csv("../raw_data/stroke_activitybyhbr.csv") %>% 
  clean_names() %>% 
  rename(health_board = "hbr")
```


```{r}
health_board_stroke <- right_join(health_board, 
                                  hb_stroke, 
                                  by = "health_board") %>% 
  select(hb_name, financial_year, admission_type, age_group, 
         sex, diagnosis, number_of_discharges, crude_rate, easr) %>%
  mutate(financial_year = str_sub(financial_year, start = 1, end = 4),
         financial_year = as.numeric(financial_year))
   # filter(is.na(hb_name))

health_board_stroke
```
